Creative Destruction in the Philosophy of Technology

### 2016-05-09 encryption/paper.tex
adwf

\section{Introduction}
Over the last few years, namely since the Snowden revelations in 2013, issues concerning privacy, anonymity and surveillence on the internet have come to the forefront of our cultural discussions.
Snowden surprised many people when he said provided evidence that the NSA has access to our internet activity, and moreover, the NSA may read our email, a mode of communication that we commonly considered to be private.
Due to this new information, many persons seeking privacy from the government resorted to using \textit{encrytion} to protect their privacy.
Encryption is means of obfuscating a message - if two people use encryption, they can talk to each other without an eavesdropper reading their message, or with extra pre-cautions, from learning their identities.

With the advent of the Snowden revelations came the slogan, ``you have nothing worry about if you have nothing to hide,'' in effect painting encryption as unpatriotic act of rebellion.

Another debate has emerged over how encryption may undermine the efforts of the law-enforcement community: encryption allows criminals to communicate, allowing them to elude law-enforcement and hence presenting a security risk to the citizenry.
Law-enforcement argues in favor of restricting the power of encryption, and in particular, requiring that all encryption used within the United States have some kind of master-key.
The master-key, which the company providing the encrpytion would hold (e.g., Apple or Google), would allow the law-enforcement agency to unlock encrypted messages.

The goal of this paper is to take a step back from the politicized arguments surrounding encryption and anonymity and ask a basic question: do we have a right to encryption?
Is ethically permissible to restrict public access to encryption, or is it unethical to give criminals easy-access to encryption?
\al{Talk about structure of paper}

\section{Background}
In order to discuss encryption from an ethical perspective, I first give background information on encryption.
Encryption has been around for centuries; we have evidence from centuries ago that various armies use some form of encryption, and more recently and famously, Hitler used the engima machine in World War II to obfuscate messages between his troops.
It is only recently that encryption has been described as an act of rebellion, and something that should potentially be restricted; my suspicion is that the main reason for this that we all know have access to encryption, as opposed to states, and so it needs to be restricted, perhaps to support a state that uses surveillence.

It is interesting, to me as a cryptographer, that we believe we can \textit{restrict} encryption.
Encryption is at its core a mathematical fact.
If I want to talk to you, we can both do some math and communicate secretly, as simple as that: math enables use to communicate privately.
To restrict encryption really means to restrict the type of encryption schemes that firms who offer encryption in their products use.
For example, Apple, who encrypts users' messages sent over iMessage, could be restricted to using an encryption scheme that uses a master-key.
If Apple does not use an encryption scheme with a master-key, a regulatory body could fine Apple for not abiding by the encryption regulations.

In contrast, the government could prevent any person from using encryption schemes with certain properties, like me using an encryption scheme that I implement myself in Python and use to communicate with my friend.
The difference between this situation and Apple in the previous paragraph is that I am not selling encrpytion; rather, I am simply doing math.
I believe that regulating encryption on the indvidiaul level is clearly a governmental overreach.
I would imagine that such regulations would be illegal by freedom of speech - the freedom to write a program - but I have been surprised by laws and court decisions before.

@@

\section{Parallel Situations}
In this section we present a variety of situations that are similar and disimilar to encryption to help us think about whether we have a fundamental right to encrytion.

First a few definitions.
Encryption, as I will be thinking about it, is a method that allows two or more parties to send messages over the internet without an eavesdropper from reading, or learning any information about, the contents of the messages.
Encryption is also used to obfuscate data on a laptop or iPhone, such as in the recent case between the FBI and Apple, but I believe that it will simplify our reasoning if we leave that situation to another day.

\al{other definitions if needed}

\subsubsection{Private Room, Private Converstation}
One way to think about encryption is verbally talking to another party in a private room, in which case we ask the question: if we are alone in a private room with another party, do we have the right to communicate privately - that is, do we have a right to \textit{not} be eavesdropped on?
To avoid annoying criticisms, let us further assume that we are in a private residence, talking in a soundproof room, and we, the conversors, have received no indication that someone might be listening to our conversation.
In this case, it seems reasonable for us to expect a private conversation - if you take issue with this, then you may ignore the following analysis.

@@

We now expand the hypothetical situation to be more similar to encryption.
Instead suppose that we are in a private residence in a crowded room. 
I am standing on one side of the room, and the person to whom I am speaking is standing on the other side.
We wish to communicate without others in the room from listening. 
We can achieve this using public-key encryption: we can each do some math, shout at each other, let everyone listen to our shouts, and in the end communicate some message that everyone else is oblivious to.

@@

The internet can be thought of as a means to communicate over a large distance in a short amount of time, like an optimized form of mail.
I write an email, send it to you via the internet, and you receive the message and open.
Since the internet runs on a public system - packets get passed around from server to server - anyone who comes across my email in the process of it getting to you could read its contents.
Therefore, the only way to communicate privately with you is to use some form of encryption.

The matter is complicated by the fact that we don't \textit{own} the internet; the internet is run by private firms and the government, and they let us use it (for a price, or taxes). 
This makes communicating over the internet similar to communicating via the United States Post Service USPS.
I give my letter to the USPS and a desired destination, the USPS does internal work to get the letter to the destination, and you eventually receive the letter.
In this framing, it feels like the USPS is doing me a favor; they are doing the hard word and heavy lifting, and I am using their convenient service.

>>>>

\section{Introduction}
Over the last few years, namely since the Snowden revelations in 2013, issues concerning privacy, anonymity and surveillence on the internet have come to the forefront of our cultural discussions.
Snowden surprised many people when he provided evidence that the NSA has access to our internet activity and specifically that the NSA may read our email, a mode of communication that we commonly considered to be private.
Due to this new information, many persons seeking privacy from the government resorted to using \textit{encrytion} to protect their privacy.
Encryption is means of obfuscating a message - if two people use encryption, they can talk to each other without worrying that an eavesdropper might be reading their messages.
In response to this, the phrase ``you have nothing worry about if you have nothing to hide'' became a slogan for those in support of the NSA; in effect, the slogan and idea it represents paints encryption as unpatriotic act of rebellion.

Another debate has emerged over how encryption may undermine the efforts of the law-enforcement community: encryption enables criminals to communicate such that law-enforcement cannot read their messages and hence presenting a national security risk.
Law-enforcement argues in favor of restricting the power of encryption, and in particular, requiring that all encryption used within the United States have some kind of master-key.
The master-key, which the company providing the encrpytion would hold (e.g., Apple or Google), would allow the law-enforcement agency to decrypt encrypted messages.

The goal of this paper is to take a step back from the politicized arguments surrounding encryption and anonymity and ask a basic question: do we have a right to encryption?
This question is difficult to answer or even approach, and as such, this paper does not answer the question, but it introduces sub-problems that need to be addressed first.
The hope was that some formalism of the encryption may come to fruition, but unfortunately, none was produced.
However, this paper is a good starting point for thinking about encryption, and is useful for future, rigorous work.

\section{Background}
In order to discuss encryption from an ethical perspective, I first give background information on encryption.
Encryption has been around for centuries; we have evidence from centuries ago that various armies use some form of encryption, and more recently and famously, Hitler used the engima machine in World War II to obfuscate messages between his troops.
It is only recently that groups believe that encryption should be restricted; my suspicion is that the main reason for this that we all know have access to encryption, as opposed to states, and so it needs to be restricted, perhaps to support a state that uses surveillence.

It is interesting to me as a cryptographer that we believe we can \textit{restrict} encryption.
Encryption is at its core a mathematical fact.
If I want to talk to you, we can both do some math and communicate secretly, as simple as that.
To restrict encryption really means to restrict the type of encryption schemes that firms who offer encryption in their products use.
For example, Apple, who encrypts users' messages sent over iMessage, could be restricted to using an encryption scheme that uses a master-key.
If Apple does not use an encryption scheme with a master-key, a regulatory body could fine Apple for not abiding by the encryption regulations.
If encryption does become regulated, I predict that some interesting black-market app will appear that performs the same essential functionality as Whatsapp.

The government can feasibly regulate the market, but they cannot regulate individual practices.
For example, it's hard to imagine the government could successfully prevent any person from using encryption schemes with certain properties, like me using an encryption scheme that I implement myself in Python and use to communicate with my friend.
The difference between this situation and Apple in the previous paragraph is that I am not selling encrpytion; rather, I am simply doing math.
I would imagine that reguations on indivudal encryption would be disallowed by freedom of speech - the freedom to write a program - but I have been surprised by laws and court decisions before.

@@

\section{Parallel Situations}
In this section we present a variety of situations that are similar and disimilar to encryption to help us think about whether we have a right to use encrytion over the internet.
Encryption, as I will be using it, is a method that allows two or more parties to send messages over the internet without an eavesdropper from reading, or learning any information about, the contents of the messages.
Encryption is also used to obfuscate data on a laptop or iPhone, such as in the recent case between the FBI and Apple, but I believe that it will simplify our reasoning if we leave that situation to another day.

\subsubsection{Private Room, Private Converstation}
If we are alone in a private room with another party, do we have the right to communicate privately - that is, do we have a right to \textit{not} be eavesdropped on?
To avoid annoying criticisms, let us further assume that we are in a private residence, talking in a soundproof room, and the conversors have received no indication that someone might be listening to our conversation.
In this case, it seems reasonable to expect a private conversation; I will operate under this assumption going forward.
However, the internet is not a private room with only two parties but a crowded space.

@@

We now expand the hypothetical situation to be more similar to encryption.
Instead suppose that we are in a private residence in a crowded room. 
One party stands on one side of the room, and the other party stands on the other side.
The parties wish to communicate without others in the room from listening.
The parties can achieve this using public-key encryption: each party can do some math, shout at each other, let everyone listen to their shouts, and in the end communicate some message that everyone else is oblivious to.

@@

The internet can be thought of as a means to communicate over a large distance in a short amount of time, like an optimized form of mail.
I write an email, send it to you via the internet, and you receive the email and read it.
Since the internet runs on a public system - packets get passed around from server to server - anyone who comes across my email in the process of it getting to you could read its contents.
Therefore, the only way to communicate privately with you is to use some form of encryption.

The matter is complicated by the fact that we don't \textit{own} the internet; the internet is run by private firms and the government, and they let us use it (for a price or taxes). 
This makes communicating over the internet similar to communicating via the United States Post Service USPS.
I give my letter to the USPS and a desired destination, the USPS does internal work to get the letter to the destination, and you eventually receive the letter.
In this framing, it feels like the USPS is doing me a favor; they are doing the hard word and heavy lifting, and I am using their convenient service.
